---
title: "TAG W3 immune cleaning"
author: "Michelle BYrne"
date: "2023-08-23"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
#knitr::opts_knit$set(root.dir = 'S:/MNHS-Psych/ANDL-Lab-TAG-Study/TAG Study/W3/saliva')

cas_dir = "/Volumes/psych-cog/dsnlab/TAG/"
proj_root = "/Volumes/psych-cog/dsnlab/TAG/projects/immune/"
data_root = "/Volumes/psych-cog/dsnlab/TAG/data/Immune/"

```

# Steps for cleaning, tidying, and reporting on inflammatory markers.
## This example is for multiplex assays of: salivary CRP(panel 1), IL-10, IL-6, TNFa (panel 2), and secretory IgA (panel 3) from Wave 3 of the University of Oregon TAG Study for a substudy K01MH111951 (michelle.byrne@gmail.com)


Calculate intra-assay CVs (from MFI, not concentration) from the sample (duplicate) CVs. Note that the duplicate sample %CVs reported in the masterfile from the lab ARE calculated from the MFI. Also, the units are PERCENTAGES even though these are not indicated (so if the value is 5.39, that is 5.39%, or 0.0539, not 539%):

```{r intra_cvs}
intraassays <- read.csv(paste0(data_root, "Wave1/TAG_W1_immune_intraassays.csv"))
                        
View(intraassays)

crp_pl1_cv <- mean(intraassays$crp_pl1)
crp_pl2_cv <- mean(intraassays$crp_pl2)
crp_pl3_cv <- mean(intraassays$crp_pl3)
crp_pl4_cv <- mean(intraassays$crp_pl4)
crp_pl5_cv <- mean(intraassays$crp_pl5)
crp_cv <- (crp_pl1_cv + crp_pl2_cv + crp_pl3_cv + crp_pl4_cv + crp_pl5_cv)/5

il10_pl1_cv <- mean(intraassays$il10_pl1)
il10_pl2_cv <- mean(intraassays$il10_pl2)
il10_pl3_cv <- mean(intraassays$il10_pl3)
il10_pl4_cv <- mean(intraassays$il10_pl4)
il10_pl5_cv <- mean(intraassays$il10_pl5)
il10_cv <- (il10_pl1_cv + il10_pl2_cv + il10_pl3_cv + il10_pl4_cv + il10_pl5_cv)/5

il6_pl1_cv <- mean(intraassays$il6_pl1)
il6_pl2_cv <- mean(intraassays$il6_pl2)
il6_pl3_cv <- mean(intraassays$il6_pl3)
il6_pl4_cv <- mean(intraassays$il6_pl4)
il6_pl5_cv <- mean(intraassays$il6_pl5)
il6_cv <- (il6_pl1_cv + il6_pl2_cv + il6_pl3_cv + il6_pl4_cv + il6_pl5_cv)/5

tnfa_pl1_cv <- mean(intraassays$tnfa_pl1)
tnfa_pl2_cv <- mean(intraassays$tnfa_pl2)
tnfa_pl3_cv <- mean(intraassays$tnfa_pl3)
tnfa_pl4_cv <- mean(intraassays$tnfa_pl4)
tnfa_pl5_cv <- mean(intraassays$tnfa_pl5)
tnfa_cv <- (tnfa_pl1_cv + tnfa_pl2_cv + tnfa_pl3_cv + tnfa_pl4_cv + tnfa_pl5_cv)/5

iga_pl1_cv <- mean(intraassays$iga_pl1)
iga_pl2_cv <- mean(intraassays$iga_pl2)
iga_pl3_cv <- mean(intraassays$iga_pl3)
iga_pl4_cv <- mean(intraassays$iga_pl4)
iga_pl5_cv <- mean(intraassays$iga_pl5)
iga_cv <- (iga_pl1_cv + iga_pl2_cv + iga_pl3_cv + iga_pl4_cv + iga_pl5_cv)/5

```

Calculate inter-assay CVs (from MFI, not concentration) from the high and low controls. Note: there are 5 plates but controls were reused for plates 4 & 5 (so only 3 plates for inter-assays) *note need to check for w3:

```{r inter_cvs}
interassays <- read_csv(paste0(data_root, "Wave1/TAG_W1_immune_interassays.csv"))
View(interassays)
# Calculate mean for controls by plate:
interassays$plate_mean <- rowMeans(interassays[,c(4:5)])

# CRP:
# Mean of high controls across plates:
crp_high_mean <- mean(data.matrix(interassays[interassays$marker=="crp"&interassays$control=="high","plate_mean"]))
# SD of high controls across plates:
crp_high_sd <- sd(data.matrix(interassays[interassays$marker=="crp"&interassays$control=="high","plate_mean"]))
# % CV high across plates:
crp_high_cv <- (crp_high_sd/crp_high_mean)*100
# Mean of low controls across plates:
crp_low_mean <- mean(data.matrix(interassays[interassays$marker=="crp"&interassays$control=="low","plate_mean"]))
# SD of low controls across plates:
crp_low_sd <- sd(data.matrix(interassays[interassays$marker=="crp"&interassays$control=="low","plate_mean"]))
# % CV low across plates:
crp_low_cv <- (crp_low_sd/crp_low_mean)*100
# AVERAGE % CV:
crp_inter_cv <- (crp_low_cv + crp_high_cv)/2

# IL-10:
# Mean of high controls across plates:
il10_high_mean <- mean(data.matrix(interassays[interassays$marker=="il10"&interassays$control=="high","plate_mean"]))
# SD of high controls across plates:
il10_high_sd <- sd(data.matrix(interassays[interassays$marker=="il10"&interassays$control=="high","plate_mean"]))
# % CV high across plates:
il10_high_cv <- (il10_high_sd/il10_high_mean)*100
# Mean of low controls across plates:
il10_low_mean <- mean(data.matrix(interassays[interassays$marker=="il10"&interassays$control=="low","plate_mean"]))
# SD of low controls across plates:
il10_low_sd <- sd(data.matrix(interassays[interassays$marker=="il10"&interassays$control=="low","plate_mean"]))
# % CV low across plates:
il10_low_cv <- (il10_low_sd/il10_low_mean)*100
# AVERAGE % CV:
il10_inter_cv <- (il10_low_cv + il10_high_cv)/2

# IL-6:
# Mean of high controls across plates:
il6_high_mean <- mean(data.matrix(interassays[interassays$marker=="il6"&interassays$control=="high","plate_mean"]))
# SD of high controls across plates:
il6_high_sd <- sd(data.matrix(interassays[interassays$marker=="il6"&interassays$control=="high","plate_mean"]))
# % CV high across plates:
il6_high_cv <- (il6_high_sd/il6_high_mean)*100
# Mean of low controls across plates:
il6_low_mean <- mean(data.matrix(interassays[interassays$marker=="il6"&interassays$control=="low","plate_mean"]))
# SD of low controls across plates:
il6_low_sd <- sd(data.matrix(interassays[interassays$marker=="il6"&interassays$control=="low","plate_mean"]))
# % CV low across plates:
il6_low_cv <- (il6_low_sd/il6_low_mean)*100
# AVERAGE % CV:
il6_inter_cv <- (il6_low_cv + il6_high_cv)/2

# TNF-a:
# Mean of high controls across plates:
tnfa_high_mean <- mean(data.matrix(interassays[interassays$marker=="tnfa"&interassays$control=="high","plate_mean"]))
# SD of high controls across plates:
tnfa_high_sd <- sd(data.matrix(interassays[interassays$marker=="tnfa"&interassays$control=="high","plate_mean"]))
# % CV high across plates:
tnfa_high_cv <- (tnfa_high_sd/tnfa_high_mean)*100
# Mean of low controls across plates:
tnfa_low_mean <- mean(data.matrix(interassays[interassays$marker=="tnfa"&interassays$control=="low","plate_mean"]))
# SD of low controls across plates:
tnfa_low_sd <- sd(data.matrix(interassays[interassays$marker=="tnfa"&interassays$control=="low","plate_mean"]))
# % CV low across plates:
tnfa_low_cv <- (tnfa_low_sd/tnfa_low_mean)*100
# AVERAGE % CV:
tnfa_inter_cv <- (tnfa_low_cv + tnfa_high_cv)/2

# SIgA (note only did high controls for this marker):
# Mean of high controls across plates:
iga_high_mean <- mean(data.matrix(interassays[interassays$marker=="iga"&interassays$control=="high","plate_mean"]))
# SD of high controls across plates:
iga_high_sd <- sd(data.matrix(interassays[interassays$marker=="iga"&interassays$control=="high","plate_mean"]))
# % CV high across plates:
iga_high_cv <- (iga_high_sd/iga_high_mean)*100
# AVERAGE % CV:
iga_inter_cv <- iga_high_cv

```

Load data and deal with out of range values (too high or low for assay to detect). 
IMPORTANT: CRP and SIgA Concentrations are in ng/ml. ALL OTHERS are in pg/ml.
We change right censored to max of distribution plus increment (i.e., sensitivity units of the assay) and left censored to lower limit of detection of that assay minus increment:
- in pg/ml, IFN-ɣ: 2.9; IL-10: 1.1; IL-12p70: 0.6; IL-2: 1.0; IL-4: 4.5; IL-6: 0.9; TNF-ɑ: 0.7
- in ng/ml: CRP: 0.001; SIgA: 0.40

Completely missing samples (i.e., no sample was given or nothing could be assayed) will be imputed after this cleaning process.

```{r}
tag_w3_immune <- read.csv(paste0(data_root, "Wave1/TAG_W1_immune_raw_concentrations.csv"))
View(tag_w3_immune)

increment = 0.01
oor_right = "OOR>"
oor_left = "OOR<"
num_oor_right_crp <- length(na.omit(tag_w3_immune$crp_raw [tag_w3_immune$crp_raw == oor_right]))
num_oor_right_il10 <- length(na.omit(tag_w3_immune$il10_raw [tag_w3_immune$il10_raw == oor_right]))
num_oor_right_il6 <- length(na.omit(tag_w3_immune$il6_raw [tag_w3_immune$il6_raw == oor_right]))
num_oor_right_tnfa <- length(na.omit(tag_w3_immune$tnfa_raw [tag_w3_immune$tnfa_raw == oor_right]))
num_oor_right_iga <- length(na.omit(tag_w3_immune$iga_raw [tag_w3_immune$iga_raw == oor_right]))

num_oor_left_crp <- length(na.omit(tag_w3_immune$crp_raw [tag_w3_immune$crp_raw == oor_left]))
num_oor_left_il10 <- length(na.omit(tag_w3_immune$il10_raw [tag_w3_immune$il10_raw == oor_left]))
num_oor_left_il6 <- length(na.omit(tag_w3_immune$il6_raw [tag_w3_immune$il6_raw == oor_left]))
num_oor_left_tnfa <- length(na.omit(tag_w3_immune$tnfa_raw [tag_w3_immune$tnfa_raw == oor_left]))
num_oor_left_iga <- length(na.omit(tag_w3_immune$iga_raw [tag_w3_immune$iga_raw == oor_left]))

# Report these in your manuscript:
num_oor_right_crp 
num_oor_right_il10
num_oor_right_il6
num_oor_right_tnfa
num_oor_right_iga
num_oor_left_crp
num_oor_left_il10
num_oor_left_il6
num_oor_left_tnfa
num_oor_left_iga

# Now replace (winsorize) the OOR samples

# First identify what the max are outside of the right censored OOR values (and force them to be numeric, because if they had OOR in them they wouldn't be):

crp_num <- tag_w3_immune[!(tag_w3_immune$crp_raw==oor_right),] 
crp_num$crp_raw <- as.numeric(crp_num$crp_raw)
max_crp <- max(na.omit(crp_num$crp_raw))

il10_num <- tag_w3_immune[!(tag_w3_immune$il10_raw==oor_right),] 
il10_num$il10_raw <- as.numeric(il10_num$il10_raw)
max_il10 <- max(na.omit(il10_num$il10_raw))

il6_num <- tag_w3_immune[!(tag_w3_immune$il6_raw==oor_right),] 
il6_num$il6_raw <- as.numeric(il6_num$il6_raw)
max_il6 <- max(na.omit(il6_num$il6_raw))

tnfa_num <- tag_w3_immune[!(tag_w3_immune$tnfa_raw==oor_right),] 
tnfa_num$tnfa_raw <- as.numeric(tnfa_num$tnfa_raw)
max_tnfa <- max(na.omit(tnfa_num$tnfa_raw))

iga_num <- tag_w3_immune[!(tag_w3_immune$iga_raw==oor_right),] 
iga_num$iga_raw <- as.numeric(iga_num$iga_raw)
max_iga <- max(na.omit(iga_num$iga_raw))

# If the immune marker is OOR> then replace with max plus increment (smaller for CRP because more sensitive))
tag_w3_immune$crp_raw <- ifelse (tag_w3_immune$crp_raw==oor_right,max_crp + (increment*.01),tag_w3_immune$crp_raw)
tag_w3_immune$il10_raw <- ifelse (tag_w3_immune$il10_raw==oor_right,max_il10 + increment,tag_w3_immune$il10_raw)
tag_w3_immune$il6_raw <- ifelse (tag_w3_immune$il6_raw==oor_right,max_il6 + increment,tag_w3_immune$il6_raw)
tag_w3_immune$tnfa_raw <- ifelse (tag_w3_immune$tnfa_raw==oor_right,max_tnfa + increment,tag_w3_immune$tnfa_raw)
tag_w3_immune$iga_raw <- ifelse (tag_w3_immune$iga_raw==oor_right,max_iga + increment,tag_w3_immune$iga_raw)

#OOR< then replace with one-half of the lower limit of detection for that assay (minus increment - smaller for CRP because more sensitive)
min_crp = 0.001/2
min_il10 = 1.1/2
min_il6 = 0.9/2
min_tnfa = 0.7/2
min_iga = 0.40/2

tag_w3_immune$crp_raw <- ifelse (tag_w3_immune$crp_raw==oor_left,min_crp - (increment*.01),tag_w3_immune$crp_raw)
tag_w3_immune$il10_raw <- ifelse (tag_w3_immune$il10_raw==oor_left,min_il10 - increment,tag_w3_immune$il10_raw)
tag_w3_immune$il6_raw <- ifelse (tag_w3_immune$il6_raw==oor_left,min_il6 - increment,tag_w3_immune$il6_raw)
tag_w3_immune$tnfa_raw <- ifelse (tag_w3_immune$tnfa_raw==oor_left,min_tnfa - increment,tag_w3_immune$tnfa_raw)
tag_w3_immune$iga_raw <- ifelse (tag_w3_immune$iga_raw==oor_left,min_iga - increment,tag_w3_immune$iga_raw)

# Then force everything to be numeric for further analyses
tag_w3_immune$crp_raw <- as.numeric(tag_w3_immune$crp_raw)
tag_w3_immune$il10_raw <- as.numeric(tag_w3_immune$il10_raw)
tag_w3_immune$il6_raw <- as.numeric(tag_w3_immune$il6_raw)
tag_w3_immune$tnfa_raw <- as.numeric(tag_w3_immune$tnfa_raw)
tag_w3_immune$iga_raw <- as.numeric(tag_w3_immune$iga_raw)

str(tag_w3_immune)

```

Correct your data for flow rate
#### Flow rate = *(Volume of saliva)/(time sample took in sec)*
#### Corrected concentration = *(observed concentration) x (Flow rate)*
##### CRP is *not* flow rate dependent
##### SIgA *is* flow rate dependent
##### For all the other markers, there is no evidence (yet) that it is flow rate dependent but not really any work done to investigate this. Will only corret for SIgA for now.

```{r}
vol_ml = 1.5
tag_w3_immune$fr <- vol_ml / tag_w3_immune$duration
tag_w3_immune$siga_frc <- tag_w3_immune$iga_raw * tag_w3_immune$fr 
# Note: will later impute siga, for now some are NA because of missing durations
```


Check normality, outliers, and skew stats in the raw concentrations
```{r check_norm}
# Histograms for immune markers to visualize normality
hist(tag_w3_immune$crp_raw)
hist(tag_w3_immune$il10_raw)
hist(tag_w3_immune$il6_raw)
hist(tag_w3_immune$tnfa_raw)
hist(tag_w3_immune$siga_frc)

#Boxplots to visualize outliers
library(ggplot2)

ggplot(tag_w3_immune, aes(x = "", y = crp_raw)) +   
  geom_boxplot() +
  ylab("CRP (ng/ml)") +
  ggtitle("TAG w3 CRP") +
  geom_smooth(method='lm', color="black")

ggplot(tag_w3_immune, aes(x = "", y = il10_raw)) +   
  geom_boxplot() +
  ylab("IL-10 (pg/ml)") +
  ggtitle("TAG w3 IL-10") +
  geom_smooth(method='lm', color="black")

ggplot(tag_w3_immune, aes(x = "", y = il6_raw)) +   
  geom_boxplot() +
  ylab("IL-6 (pg/ml)") +
  ggtitle("TAG w3 IL-6") +
  geom_smooth(method='lm', color="black")

ggplot(tag_w3_immune, aes(x = "", y = tnfa_raw)) +   
  geom_boxplot() +
  ylab("TNF-a (pg/ml)") +
  ggtitle("TAG w3 TNF-a") +
  geom_smooth(method='lm', color="black")

ggplot(tag_w3_immune, aes(x = "", y = siga_frc)) +   
  geom_boxplot() +
  ylab("SIgA (ng/ml)") +
  ggtitle("TAG w3 SIgA") +
  geom_smooth(method='lm', color="black")

#  Calculate and report normality statistics (skew and kurtosis)
# Kurtosis and Skew should be -2/+2 (West, et al. 1995)
library(psych)
crp_summary <- describe(tag_w3_immune$crp_raw)
crp_summary
il10_summary <- describe(tag_w3_immune$il10_raw)
il10_summary
il6_summary <- describe(tag_w3_immune$il6_raw)
il6_summary
tnfa_summary <- describe(tag_w3_immune$tnfa_raw)
tnfa_summary
siga_summary <- describe(tag_w3_immune$siga_frc)
siga_summary
```

If necessary, transform data

```{r}
 
# if skew or kurtosis is < -2 or > 2 then calculate the natural log of that variable
#Note: although it is common to do a natural log or square root transformation, these transfromations did not result in acceptable skew/kurtosis stats. Therefore, we explored also using the transform Tukey function, which uses iterative Shapiro-Wilk tests, and finds the lambda values that maximizes the W statistic from those tests. (e.g., find power trasnformation that makes the data fit the normal distribution as closely as possible). See info on rcompanion for more. We hae provided both natural log (ln_) output variables and tukey transformed (tu_)

library(rcompanion)

if (crp_summary$skew > 2 | crp_summary$skew < -2 | crp_summary$kurtosis > 2 | crp_summary$kurtosis < -2) {
  tag_w3_immune$ln_crp <- log(tag_w3_immune$crp_raw)
}

if (il10_summary$skew > 2 | il10_summary$skew < -2 | il10_summary$kurtosis > 2 | il10_summary$kurtosis < -2) {
  tag_w3_immune$ln_il10 <- log(tag_w3_immune$il10_raw)
}

if (il6_summary$skew > 2 | il6_summary$skew < -2 | il6_summary$kurtosis > 2 | il6_summary$kurtosis < -2) {
  tag_w3_immune$ln_il6 <- log(tag_w3_immune$il6_raw)
}

if (tnfa_summary$skew > 2 | tnfa_summary$skew < -2 | tnfa_summary$kurtosis > 2 | tnfa_summary$kurtosis < -2) {
  tag_w3_immune$ln_tnfa <- log(tag_w3_immune$tnfa_raw)
}

if (siga_summary$skew > 2 | siga_summary$skew < -2 | siga_summary$kurtosis > 2 | siga_summary$kurtosis < -2) {
  tag_w3_immune$ln_siga <- log(tag_w3_immune$siga_frc)
}

# Tukey:

if (crp_summary$skew > 2 | crp_summary$skew < -2 | crp_summary$kurtosis > 2 | crp_summary$kurtosis < -2) {
  tag_w3_immune$tu_crp <- transformTukey(tag_w3_immune$crp_raw)
}

if (il10_summary$skew > 2 | il10_summary$skew < -2 | il10_summary$kurtosis > 2 | il10_summary$kurtosis < -2) {
  tag_w3_immune$tu_il10 <- transformTukey(tag_w3_immune$il10_raw)
}

if (il6_summary$skew > 2 | il6_summary$skew < -2 | il6_summary$kurtosis > 2 | il6_summary$kurtosis < -2) {
  tag_w3_immune$tu_il6 <- transformTukey(tag_w3_immune$il6_raw)
}

if (tnfa_summary$skew > 2 | tnfa_summary$skew < -2 | tnfa_summary$kurtosis > 2 | tnfa_summary$kurtosis < -2) {
  tag_w3_immune$tu_tnfa <- transformTukey(tag_w3_immune$tnfa_raw)
}

if (siga_summary$skew > 2 | siga_summary$skew < -2 | siga_summary$kurtosis > 2 | siga_summary$kurtosis < -2) {
  tag_w3_immune$tu_siga <- transformTukey(tag_w3_immune$siga_frc)
}

# Check everything post transformation for any transformed variables 
final_lncrp_summary <- describe(tag_w3_immune$ln_crp)
final_lncrp_summary
final_lnil10_summary <- describe(tag_w3_immune$ln_il10)
final_lnil10_summary
final_lnil6_summary <- describe(tag_w3_immune$ln_il6)
final_lnil6_summary 
final_lntnfa_summary <- describe(tag_w3_immune$ln_tnfa)
final_lntnfa_summary
final_lnsiga_summary <- describe(tag_w3_immune$ln_siga)
final_lnsiga_summary

final_tucrp_summary <- describe(tag_w3_immune$tu_crp)
final_tucrp_summary
final_tuil10_summary <- describe(tag_w3_immune$tu_il10)
final_tuil10_summary
final_tuil6_summary <- describe(tag_w3_immune$tu_il6)
final_tuil6_summary 
final_tutnfa_summary <- describe(tag_w3_immune$tu_tnfa)
final_tutnfa_summary
final_tusiga_summary <- describe(tag_w3_immune$tu_siga)
final_tusiga_summary

hist(tag_w3_immune$ln_crp)
hist(tag_w3_immune$ln_il10)
hist(tag_w3_immune$ln_il6)
hist(tag_w3_immune$ln_tnfa)
hist(tag_w3_immune$ln_siga)

hist(tag_w3_immune$tu_crp)
hist(tag_w3_immune$tu_il10)
hist(tag_w3_immune$tu_il6)
hist(tag_w3_immune$tu_tnfa)
hist(tag_w3_immune$tu_siga)

ggplot(tag_w3_immune, aes(x = "", y = ln_crp)) +   
  geom_boxplot() +
  ylab("log (CRP)") +
  ggtitle("TAG w3 CRP (ln transformed)") +
  geom_smooth(method='lm', color="black")
ggplot(tag_w3_immune, aes(x = "", y = ln_il10)) +   
  geom_boxplot() +
  ylab("log (IL-10)") +
  ggtitle("TAG w3 IL-10 (ln transformed)") +
  geom_smooth(method='lm', color="black")
ggplot(tag_w3_immune, aes(x = "", y = ln_il6)) +   
  geom_boxplot() +
  ylab("log (IL-6)") +
  ggtitle("TAG w3 IL-6 (ln transformed)") +
  geom_smooth(method='lm', color="black")
ggplot(tag_w3_immune, aes(x = "", y = ln_tnfa)) +   
  geom_boxplot() +
  ylab("log (TNF-a)") +
  ggtitle("TAG w3 TNF-a (ln transformed)") +
  geom_smooth(method='lm', color="black")
ggplot(tag_w3_immune, aes(x = "", y = ln_siga)) +   
  geom_boxplot() +
  ylab("log (SIgA)") +
  ggtitle("TAG w3 SIgA (ln transformed)") +
  geom_smooth(method='lm', color="black")

# tukey box plots:
ggplot(tag_w3_immune, aes(x = "", y = tu_crp)) +   
  geom_boxplot() +
  ylab("log (CRP)") +
  ggtitle("TAG w3 CRP (tu transformed)") +
  geom_smooth(method='lm', color="black")
ggplot(tag_w3_immune, aes(x = "", y = tu_il10)) +   
  geom_boxplot() +
  ylab("log (IL-10)") +
  ggtitle("TAG w3 IL-10 (tu transformed)") +
  geom_smooth(method='lm', color="black")
ggplot(tag_w3_immune, aes(x = "", y = tu_il6)) +   
  geom_boxplot() +
  ylab("log (IL-6)") +
  ggtitle("TAG w3 IL-6 (tu transformed)") +
  geom_smooth(method='lm', color="black")
ggplot(tag_w3_immune, aes(x = "", y = tu_tnfa)) +   
  geom_boxplot() +
  ylab("log (TNF-a)") +
  ggtitle("TAG w3 TNF-a (tu transformed)") +
  geom_smooth(method='lm', color="black")
ggplot(tag_w3_immune, aes(x = "", y = tu_siga)) +   
  geom_boxplot() +
  ylab("log (SIgA)") +
  ggtitle("TAG w3 SIgA (tu transformed)") +
  geom_smooth(method='lm', color="black")
```

Winsorize any remaining outliers that are +/- 3 SD (if any)

```{r}
lncrp_up_limit <- (final_lncrp_summary$mean + 3*(final_lncrp_summary$sd))
lncrp_up_limit
lncrp_lo_limit <- (final_lncrp_summary$mean - 3*(final_lncrp_summary$sd))
lncrp_lo_limit

lnil10_up_limit <- (final_lnil10_summary$mean + 3*(final_lnil10_summary$sd))
lnil10_up_limit
lnil10_lo_limit <- (final_lnil10_summary$mean - 3*(final_lnil10_summary$sd))
lnil10_lo_limit

lnil6_up_limit <- (final_lnil6_summary$mean + 3*(final_lnil6_summary$sd))
lnil6_up_limit
lnil6_lo_limit <- (final_lnil6_summary$mean - 3*(final_lnil6_summary$sd))
lnil6_lo_limit

lntnfa_up_limit <- (final_lntnfa_summary$mean + 3*(final_lntnfa_summary$sd))
lntnfa_up_limit
lntnfa_lo_limit <- (final_lntnfa_summary$mean - 3*(final_lntnfa_summary$sd))
lntnfa_lo_limit

lnsiga_up_limit <- (final_lnsiga_summary$mean + 3*(final_lnsiga_summary$sd))
lnsiga_up_limit
lnsiga_lo_limit <- (final_lnsiga_summary$mean - 3*(final_lnsiga_summary$sd))
lnsiga_lo_limit


tucrp_up_limit <- (final_tucrp_summary$mean + 3*(final_tucrp_summary$sd))
tucrp_up_limit
tucrp_lo_limit <- (final_tucrp_summary$mean - 3*(final_tucrp_summary$sd))
tucrp_lo_limit

tuil10_up_limit <- (final_tuil10_summary$mean + 3*(final_tuil10_summary$sd))
tuil10_up_limit
tuil10_lo_limit <- (final_tuil10_summary$mean - 3*(final_tuil10_summary$sd))
tuil10_lo_limit

tuil6_up_limit <- (final_tuil6_summary$mean + 3*(final_tuil6_summary$sd))
tuil6_up_limit
tuil6_lo_limit <- (final_tuil6_summary$mean - 3*(final_tuil6_summary$sd))
tuil6_lo_limit

tutnfa_up_limit <- (final_tutnfa_summary$mean + 3*(final_tutnfa_summary$sd))
tutnfa_up_limit
tutnfa_lo_limit <- (final_tutnfa_summary$mean - 3*(final_tutnfa_summary$sd))
tutnfa_lo_limit

tusiga_up_limit <- (final_tusiga_summary$mean + 3*(final_tusiga_summary$sd))
tusiga_up_limit
tusiga_lo_limit <- (final_tusiga_summary$mean - 3*(final_tusiga_summary$sd))
tusiga_lo_limit

# First check if there are any outliers (if not, no need to winsorize). Also, report these numbers. 
lncrp_num_right_outliers <- length(na.omit(tag_w3_immune$ln_crp [tag_w3_immune$ln_crp > lncrp_up_limit])) 
lncrp_num_right_outliers
lncrp_num_left_outliers <- length(na.omit(tag_w3_immune$ln_crp [tag_w3_immune$ln_crp < lncrp_lo_limit]))
lncrp_num_left_outliers

lnil10_num_right_outliers <- length(na.omit(tag_w3_immune$ln_il10 [tag_w3_immune$ln_il10 > lnil10_up_limit])) 
lnil10_num_right_outliers
lnil10_num_left_outliers <- length(na.omit(tag_w3_immune$ln_il10 [tag_w3_immune$ln_il10 < lnil10_lo_limit])) 
lnil10_num_left_outliers

lnil6_num_right_outliers <- length(na.omit(tag_w3_immune$ln_il6 [tag_w3_immune$ln_il6 > lnil6_up_limit])) 
lnil6_num_right_outliers
lnil6_num_left_outliers <- length(na.omit(tag_w3_immune$ln_il6 [tag_w3_immune$ln_il6 < lnil6_lo_limit]))
lnil6_num_left_outliers

lntnfa_num_right_outliers <- length(na.omit(tag_w3_immune$ln_tnfa [tag_w3_immune$ln_tnfa > lntnfa_up_limit])) 
lntnfa_num_right_outliers
lntnfa_num_left_outliers <- length(na.omit(tag_w3_immune$ln_tnfa [tag_w3_immune$ln_tnfa < lntnfa_lo_limit])) 
lntnfa_num_left_outliers 

lnsiga_num_right_outliers <- length(na.omit(tag_w3_immune$ln_siga [tag_w3_immune$ln_siga > lnsiga_up_limit]))
lnsiga_num_right_outliers 
lnsiga_num_left_outliers <- length(na.omit(tag_w3_immune$ln_siga [tag_w3_immune$ln_siga < lnsiga_lo_limit]))
lnsiga_num_left_outliers

tucrp_num_right_outliers <- length(na.omit(tag_w3_immune$tu_crp [tag_w3_immune$tu_crp > tucrp_up_limit])) 
tucrp_num_right_outliers
tucrp_num_left_outliers <- length(na.omit(tag_w3_immune$tu_crp [tag_w3_immune$tu_crp < tucrp_lo_limit]))
tucrp_num_left_outliers

tuil10_num_right_outliers <- length(na.omit(tag_w3_immune$tu_il10 [tag_w3_immune$tu_il10 > tuil10_up_limit])) 
tuil10_num_right_outliers
tuil10_num_left_outliers <- length(na.omit(tag_w3_immune$tu_il10 [tag_w3_immune$tu_il10 < tuil10_lo_limit])) 
tuil10_num_left_outliers

tuil6_num_right_outliers <- length(na.omit(tag_w3_immune$tu_il6 [tag_w3_immune$tu_il6 > tuil6_up_limit])) 
tuil6_num_right_outliers
tuil6_num_left_outliers <- length(na.omit(tag_w3_immune$tu_il6 [tag_w3_immune$tu_il6 < tuil6_lo_limit]))
tuil6_num_left_outliers

tutnfa_num_right_outliers <- length(na.omit(tag_w3_immune$tu_tnfa [tag_w3_immune$tu_tnfa > tutnfa_up_limit])) 
tutnfa_num_right_outliers
tutnfa_num_left_outliers <- length(na.omit(tag_w3_immune$tu_tnfa [tag_w3_immune$tu_tnfa < tutnfa_lo_limit])) 
tutnfa_num_left_outliers 

tusiga_num_right_outliers <- length(na.omit(tag_w3_immune$tu_siga [tag_w3_immune$tu_siga > tusiga_up_limit]))
tusiga_num_right_outliers 
tusiga_num_left_outliers <- length(na.omit(tag_w3_immune$tu_siga [tag_w3_immune$tu_siga < tusiga_lo_limit]))
tusiga_num_left_outliers
```

* Note: Update for W3. (From W1: In this data, tukey has 1 left outlier for IFN-g and 1 left outlier for IL-10. For ln, 6 right IFN-g, 1 left IL-10, 3 right IL-12, 2 right IL-4, 1 right TNF-a. Winsorize those.)
```{r}
library(DescTools)
tag_w3_immune$ifng_tu_w <- Winsorize(tag_w3_immune$tu_ifng, minval = tuifng_lo_limit, maxval = tuifng_up_limit, na.rm = TRUE)
tag_w3_immune$il10_tu_w <- Winsorize(tag_w3_immune$tu_il10, minval = tuil10_lo_limit, maxval = tuil10_up_limit, na.rm = TRUE)

tag_w3_immune$ifng_ln_w <- Winsorize(tag_w3_immune$ln_ifng, minval = lnifng_lo_limit, maxval = lnifng_up_limit, na.rm = TRUE)
tag_w3_immune$il10_ln_w <- Winsorize(tag_w3_immune$ln_il10, minval = lnil10_lo_limit, maxval = lnil10_up_limit, na.rm = TRUE)
tag_w3_immune$il12_ln_w <- Winsorize(tag_w3_immune$ln_il12, minval = lnil12_lo_limit, maxval = lnil12_up_limit, na.rm = TRUE)
tag_w3_immune$il4_ln_w <- Winsorize(tag_w3_immune$ln_il4, minval = lnil4_lo_limit, maxval = lnil4_up_limit, na.rm = TRUE)
tag_w3_immune$tnfa_ln_w <- Winsorize(tag_w3_immune$ln_tnfa, minval = lntnfa_lo_limit, maxval = lntnfa_up_limit, na.rm = TRUE)
```


Output final cleaned data
```{r}
tag_w3_immune$crp_tu_cleaned <- tag_w3_immune$tu_crp
tag_w3_immune$il10_tu_cleaned <- tag_w3_immune$il10_tu_w
tag_w3_immune$il6_tu_cleaned <- tag_w3_immune$tu_il6
tag_w3_immune$tnfa_tu_cleaned <- tag_w3_immune$tu_tnfa
tag_w3_immune$siga_tu_cleaned <- tag_w3_immune$tu_siga

tag_w3_immune$crp_ln_cleaned <- tag_w3_immune$ln_crp
tag_w3_immune$il10_ln_cleaned <- tag_w3_immune$il10_ln_w
tag_w3_immune$il6_ln_cleaned <- tag_w3_immune$ln_il6
tag_w3_immune$tnfa_ln_cleaned <- tag_w3_immune$tnfa_ln_w
tag_w3_immune$siga_ln_cleaned <- tag_w3_immune$ln_siga

write.csv(tag_w3_immune, file="TAG_W3_immune_cleaned.csv")
```

