---
title: "SOS Bioreactivity Analyses"
date: today
format:
  html:
    toc: true
    toc-float: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show code"
    theme: journal
    embed-resources: true
execute:
  warning: false
  message: false
---

```{r setup, echo=FALSE}
pacman::p_load(
  "dplyr", "ggplot2", "tidyr", "stringr", "purrr", "here",
  "lme4", "lmerTest", "broom", "broom.mixed", "readxl",
  "ggpubr", "ggsignif", "Hmisc", "corrplot", "knitr", "kableExtra", "ggdist"
)

proj_path <- here()
data_path <- Sys.getenv("CAS_DIR")
options(knitr.table.format = "html")
```

```{r load-data, echo=FALSE}
sos_cytokines <- read.csv(file.path(
  data_path, "TAG/data/SOS/saliva/Cleaned_datasets/SOS_cytokines_cleaned.csv"
))
sos_crp <- read.csv(file.path(
  data_path, "TAG/data/SOS/saliva/Cleaned_datasets/SOS_CRP_cleaned.csv"
))
sos_hormones <- read.csv(file.path(
  data_path, "TAG/data/SOS/saliva/Cleaned_datasets/SOS_hormones_cleaned.csv"
)) |>
  select(-starts_with("Estradiol"))

hr_resting1_raw  <- read.csv(file.path(data_path, "TAG/data/SOS/physio/Mean_HR_Resting1.csv"))
hr_resting2_raw  <- read.csv(file.path(data_path, "TAG/data/SOS/physio/Mean_HR_Resting2.csv"))
hr_task_raw      <- read.csv(file.path(data_path, "TAG/data/SOS/physio/Mean_HR_Task.csv"))
hr_conditions_raw <- read_excel(file.path(
  data_path, "TAG/data/SOS/physio/Heart_Rate_Data_Task_Sequence_All_Slices_Conditions.xlsx"
))
```

```{r hr-processing, echo=FALSE}
hr_exclude <- c(
  "TAG313p2_20191117_151413",   # duplicate session
  "TAG077_20190823_145422",     # artifact: impossibly low task HR
  "TAG188_20190731_145115"      # artifact: impossibly low task HR
)

hr_resting1 <- hr_resting1_raw |> rename(id = 1) |> filter(!id %in% hr_exclude) |> mutate(sequence = "Resting 1")
hr_resting2 <- hr_resting2_raw |> rename(id = 1) |> filter(!id %in% hr_exclude) |> mutate(sequence = "Resting 2")
hr_task      <- hr_task_raw    |> rename(id = 1) |> filter(!id %in% hr_exclude) |> mutate(sequence = "Task")

hr_all_wide <- bind_rows(hr_resting1, hr_resting2, hr_task) |>
  pivot_wider(names_from = sequence, values_from = Mean_HR) |>
  mutate(resting_avg = if_else(is.na(`Resting 2`), `Resting 1`, (`Resting 1` + `Resting 2`) / 2))

hr_conditions_avg <- hr_conditions_raw |>
  pivot_longer(SOS068_20190303_150846:TAG223_20191116_145342, names_to = "id", values_to = "hr") |>
  rename(task_condition = Task_condition) |>
  group_by(id, task_condition) |>
  summarise(mean_hr = mean(hr, na.rm = TRUE), .groups = "drop") |>
  filter(!is.na(task_condition), task_condition != "NA") |>
  mutate(task_condition = factor(task_condition,
    levels = c("1", "2", "3", "R"),
    labels = c("Other", "Self-Disconnected", "Self-Connected", "Rest")
  )) |>
  filter(!id %in% hr_exclude)

pull_hr <- function(cond) filter(hr_conditions_avg, task_condition == cond)$mean_hr

make_box_stats <- function(data, col) {
  function(y, upper_limit = max(data[[col]], na.rm = TRUE) * 1.15) {
    data.frame(
      y     = 0.95 * upper_limit,
      label = paste("Mean =", round(mean(y), 2), "\n", "Median =", round(median(y), 2), "\n")
    )
  }
}
```

```{r hr-difference, echo=FALSE}
hr_final_wide <- hr_all_wide |>
  full_join(
    hr_conditions_avg |> pivot_wider(names_from = task_condition, values_from = mean_hr),
    by = "id"
  ) |>
  mutate(
    SampleID = case_when(
      str_sub(id, 1, 6) == "SOS068" ~ "TAG068",
      str_sub(id, 1, 6) == "SOS076" ~ "TAG076",
      .default = str_sub(id, 1, 6)
    ),
    task_rs_hr_diff    = Task - resting_avg,
    conn_dis_hr_diff   = `Self-Connected` - `Self-Disconnected`,
    conn_other_hr_diff = `Self-Connected` - Other,
    conn_rest_hr_diff  = `Self-Connected` - Rest
  )

hr_difference <- hr_final_wide |>
  select(SampleID, task_rs_hr_diff, conn_dis_hr_diff, conn_other_hr_diff, conn_rest_hr_diff)
```

```{r biomarker-processing, echo=FALSE}
sos_crp_panel       <- sos_crp       |> select(SampleID, Time, CRP_log)
sos_cytokines_panel <- sos_cytokines |> select(SampleID, Time, IL10_log, IL6_log, TNFalpha_log)
sos_hormones_panel  <- sos_hormones  |> select(SampleID, Time, DHEA_log, Testosterone_log, Cortisol_log)

run_mlm <- function(outcome, data) {
  f_lin  <- as.formula(paste(outcome, "~ 1 + Time + (1 + Time | SampleID)"))
  f_quad <- as.formula(paste(outcome, "~ 1 + poly(Time, 2) + (1 + poly(Time, 2) | SampleID)"))
  mlm_lin  <- lmer(f_lin,  na.action = na.omit, REML = TRUE, data = data)
  mlm_quad <- lmer(f_quad, na.action = na.omit, REML = TRUE, data = data)
  list(linear = mlm_lin, quadratic = mlm_quad, aic = AIC(mlm_lin, mlm_quad), bic = BIC(mlm_lin, mlm_quad))
}

mlm_crp      <- run_mlm("CRP_log",          sos_crp_panel)
mlm_il10     <- run_mlm("IL10_log",         sos_cytokines_panel)
mlm_il6      <- run_mlm("IL6_log",          sos_cytokines_panel)
mlm_tnfalpha <- run_mlm("TNFalpha_log",     sos_cytokines_panel)
mlm_dhea     <- run_mlm("DHEA_log",         sos_hormones_panel)
mlm_test     <- run_mlm("Testosterone_log", sos_hormones_panel)
mlm_cort     <- run_mlm("Cortisol_log",     sos_hormones_panel)
```

```{r peaks-processing, echo=FALSE}
sos_crp_wide <- sos_crp_panel |>
  pivot_wider(names_from = Time, values_from = CRP_log, names_glue = "CRP_log_{Time}", values_fn = mean) |>
  mutate(crp_peak = if_else(CRP_log_3 > CRP_log_4, CRP_log_3, CRP_log_4))

sos_cytokines_wide <- sos_cytokines_panel |>
  pivot_wider(names_from = Time, values_from = c(IL10_log, IL6_log, TNFalpha_log), values_fn = mean) |>
  mutate(
    il10_peak = if_else(IL10_log_3     > IL10_log_4,     IL10_log_3,     IL10_log_4),
    il6_peak  = if_else(IL6_log_3      > IL6_log_4,      IL6_log_3,      IL6_log_4),
    tnfa_peak = if_else(TNFalpha_log_3 > TNFalpha_log_4, TNFalpha_log_3, TNFalpha_log_4)
  )

sos_hormones_wide <- sos_hormones_panel |>
  pivot_wider(names_from = Time, values_from = c(DHEA_log, Testosterone_log, Cortisol_log), values_fn = mean) |>
  mutate(
    dhea_peak = if_else(DHEA_log_3         > DHEA_log_4,         DHEA_log_3,         DHEA_log_4),
    test_peak = if_else(Testosterone_log_3 > Testosterone_log_4, Testosterone_log_3, Testosterone_log_4),
    cort_peak = if_else(Cortisol_log_3     > Cortisol_log_4,     Cortisol_log_3,     Cortisol_log_4)
  )

make_peak_long <- function(wide_data, t1_col, peak_col, value_name) {
  wide_data |>
    select(SampleID, all_of(c(t1_col, peak_col))) |>
    pivot_longer(all_of(c(t1_col, peak_col)), names_to = "time_point", values_to = value_name) |>
    mutate(time_point = factor(time_point, levels = c(t1_col, peak_col), labels = c("T1", "T3/T4 peak")))
}

fmt_p_fdr <- function(p) {
  if (p < 0.001) "p_FDR < 0.001" else paste0("p_FDR = ", round(p, 3))
}

biomarker_plot <- function(data, y_col, y_label, p_fdr) {
  ggplot(data, aes(x = time_point, y = .data[[y_col]])) +
    stat_halfeye(
      aes(fill = time_point),
      adjust = 0.5, width = 0.6, .width = 0,
      justification = -0.2, point_colour = NA, alpha = 0.8
    ) +
    geom_boxplot(width = 0.15, outlier.shape = NA, color = "gray30", linewidth = 0.5) +
    geom_point(
      aes(color = time_point),
      position = position_jitter(seed = 1, width = 0.1),
      size = 1.2, alpha = 0.3
    ) +
    scale_fill_viridis_d() +
    scale_color_viridis_d(guide = "none") +
    geom_signif(
      comparisons = list(c("T1", "T3/T4 peak")),
      annotations = fmt_p_fdr(p_fdr),
      textsize = 3.2, vjust = -0.2
    ) +
    ylab(y_label) + xlab(NULL) + theme_minimal() + theme(legend.position = "none")
}

sos_crp_long  <- make_peak_long(sos_crp_wide,       "CRP_log_1",          "crp_peak",  "crp_log_conc")
sos_il10_long <- make_peak_long(sos_cytokines_wide,  "IL10_log_1",         "il10_peak", "il10_log_conc")
sos_il6_long  <- make_peak_long(sos_cytokines_wide,  "IL6_log_1",          "il6_peak",  "il6_log_conc")
sos_tnfa_long <- make_peak_long(sos_cytokines_wide,  "TNFalpha_log_1",     "tnfa_peak", "tnfa_log_conc")
sos_dhea_long <- make_peak_long(sos_hormones_wide,   "DHEA_log_1",         "dhea_peak", "dhea_log_conc")
sos_test_long <- make_peak_long(sos_hormones_wide,   "Testosterone_log_1", "test_peak", "test_log_conc")
sos_cort_long <- make_peak_long(sos_hormones_wide,   "Cortisol_log_1",     "cort_peak", "cort_log_conc")

run_peak_test <- function(peak_col, t1_col, data) {
  t.test(data[[peak_col]], data[[t1_col]], alternative = "greater", paired = TRUE)
}

crp_difference <- sos_crp_wide |>
  mutate(crp_difference = crp_peak - CRP_log_1) |>
  select(SampleID, crp_difference)

cytokines_difference <- sos_cytokines_wide |>
  mutate(
    il10_difference = il10_peak - IL10_log_1,
    il6_difference  = il6_peak  - IL6_log_1,
    tnfa_difference = tnfa_peak - TNFalpha_log_1
  ) |>
  select(SampleID, il10_difference, il6_difference, tnfa_difference)

hormones_difference <- sos_hormones_wide |>
  mutate(
    dhea_difference = dhea_peak - DHEA_log_1,
    test_difference = test_peak - Testosterone_log_1,
    cort_difference = cort_peak - Cortisol_log_1
  ) |>
  select(SampleID, dhea_difference, test_difference, cort_difference)

biomarkers_difference <- crp_difference |>
  full_join(cytokines_difference, by = "SampleID") |>
  full_join(hormones_difference,  by = "SampleID")

bio_all_difference <- hr_difference |>
  full_join(biomarkers_difference, by = "SampleID")

# Subjective stress data not yet available; will be added when CSV is accessible
all_difference <- bio_all_difference

# Peak t-tests and FDR correction (immune and hormones corrected separately)
peak_tests_raw <- list(
  crp  = run_peak_test("crp_peak",  "CRP_log_1",          sos_crp_wide),
  il10 = run_peak_test("il10_peak", "IL10_log_1",         sos_cytokines_wide),
  il6  = run_peak_test("il6_peak",  "IL6_log_1",          sos_cytokines_wide),
  tnfa = run_peak_test("tnfa_peak", "TNFalpha_log_1",     sos_cytokines_wide),
  dhea = run_peak_test("dhea_peak", "DHEA_log_1",         sos_hormones_wide),
  test = run_peak_test("test_peak", "Testosterone_log_1", sos_hormones_wide),
  cort = run_peak_test("cort_peak", "Cortisol_log_1",     sos_hormones_wide)
)
peak_praw <- map_dbl(peak_tests_raw, \(t) t$p.value)
peak_pfdr <- setNames(
  c(p.adjust(peak_praw[1:4], method = "fdr"),
    p.adjust(peak_praw[5:7], method = "fdr")),
  names(peak_praw)
)
```

---

# Heart Rate

## Descriptive Means

```{r hr-descriptives}
tibble(
  Condition = c(
    "Resting 1", "Resting 2", "Resting Average", "Task (overall)",
    "Within-task: Rest", "Within-task: Other",
    "Within-task: Self-Disconnected", "Within-task: Self-Connected"
  ),
  `Mean HR (bpm)` = round(c(
    mean(hr_all_wide$`Resting 1`),
    mean(hr_all_wide$`Resting 2`, na.rm = TRUE),
    mean(hr_all_wide$resting_avg),
    mean(hr_all_wide$Task),
    mean(pull_hr("Rest")),
    mean(pull_hr("Other")),
    mean(pull_hr("Self-Disconnected")),
    mean(pull_hr("Self-Connected"))
  ), 2)
) |>
  kable()
```

## HR Difference Score Descriptives

```{r hr-diff-descriptives}
bind_rows(
  hr_difference |> select(-SampleID) |> summarise(across(everything(), \(x) mean(x, na.rm = TRUE))) |> mutate(stat = "Mean"),
  hr_difference |> select(-SampleID) |> summarise(across(everything(), \(x) sd(x,   na.rm = TRUE))) |> mutate(stat = "SD")
) |>
  select(stat, everything()) |>
  rename(
    `ΔHR: Task vs. Rest`  = task_rs_hr_diff,
    `ΔHR: SC vs. SD`      = conn_dis_hr_diff,
    `ΔHR: SC vs. Other`   = conn_other_hr_diff,
    `ΔHR: SC vs. Rest`    = conn_rest_hr_diff
  ) |>
  kable(digits = 3, caption = "HR reactivity difference scores (positive = higher during more self-connected/active condition)")
```

## Paired t-tests

```{r hr-ttests}
hr_ttest_tbl <- list(
  "Task vs. Resting" = t.test(hr_all_wide$Task, hr_all_wide$resting_avg, alternative = "greater", paired = TRUE),
  "Self-Connected vs. Rest" = t.test(pull_hr("Self-Connected"), pull_hr("Rest"), alternative = "greater", paired = TRUE),
  "Self-Connected vs. Self-Disconnected" = t.test(pull_hr("Self-Connected"), pull_hr("Self-Disconnected"), alternative = "greater", paired = TRUE),
  "Self-Connected vs. Other" = t.test(pull_hr("Self-Connected"), pull_hr("Other"), alternative = "greater", paired = TRUE)
) |>
  map(broom::tidy) |>
  list_rbind(names_to = "comparison") |>
  select(comparison, estimate, statistic, parameter, p.value) |>
  mutate(across(where(is.numeric), \(x) round(x, 3)))

hr_ttest_tbl |>
  mutate(p.value = cell_spec(p.value, bold = p.value < 0.05,
    color = if_else(p.value < 0.05, "#2166ac", "black"))) |>
  kable(col.names = c("Comparison", "Mean diff", "t", "df", "p"), escape = FALSE) |>
  kable_styling(bootstrap_options = "hover", full_width = FALSE)
```

## Plots {.tabset}

### Task vs. Resting

```{r avg-hr-plot, fig.width=5, fig.height=5}
hr_avg_long <- hr_all_wide |>
  select(id, resting_avg, Task) |>
  pivot_longer(
    c(resting_avg, Task),
    names_to = "condition",
    values_to = "mean_hr"
  ) |>
  mutate(
    condition = factor(
      condition,
      levels = c("resting_avg", "Task"),
      labels = c("Resting State Scan", "SOS Task")
    )
  )

ggplot(hr_avg_long, aes(x = condition, y = mean_hr)) +
  stat_halfeye(
    aes(fill = condition),
    adjust = 0.5,
    width = 0.6,
    .width = 0,
    justification = -0.2,
    point_colour = NA,
    alpha = 0.8
  ) +
  geom_boxplot(
    width = 0.15,
    outlier.shape = NA,
    color = "gray30",
    linewidth = 0.5
  ) +
  geom_point(
    aes(color = condition),
    position = position_jitter(seed = 1, width = 0.1),
    size = 1.2,
    alpha = 0.3
  ) +
  scale_fill_viridis_d() +
  scale_color_viridis_d(guide = "none") +
  stat_compare_means(
    method = "t.test",
    paired = TRUE,
    comparisons = list(c("Resting State Scan", "SOS Task")),
    method.args = list(alternative = "greater")
  ) +
  ylab("Mean Heart Rate (bpm)") +
  xlab(NULL) +
  theme_minimal() +
  theme(legend.position = "none")
```

### Within-task Conditions

```{r task-hr-plot, fig.width=7, fig.height=5}
task_hr_comparisons <- list(
  c("Self-Connected", "Rest"), c("Self-Connected", "Self-Disconnected"),
  c("Self-Connected", "Other"), c("Self-Disconnected", "Other"),
  c("Self-Disconnected", "Rest"), c("Other", "Rest")
)

ggplot(hr_conditions_avg, aes(x = task_condition, y = mean_hr)) +
  stat_halfeye(
    aes(fill = task_condition),
    adjust = 0.5, width = 0.6, .width = 0,
    justification = -0.2, point_colour = NA, alpha = 0.8
  ) +
  geom_boxplot(width = 0.15, outlier.shape = NA, color = "gray30", linewidth = 0.5) +
  geom_point(
    aes(color = task_condition),
    position = position_jitter(seed = 1, width = 0.1),
    size = 1.2, alpha = 0.3
  ) +
  scale_fill_viridis_d() +
  scale_color_viridis_d(guide = "none") +
  stat_compare_means(
    method = "t.test", paired = TRUE, comparisons = task_hr_comparisons,
    method.args = list(alternative = "greater")
  ) +
  ylab("Mean Heart Rate (bpm)") + xlab(NULL) + theme_minimal() + theme(legend.position = "none")
```

---

# Salivary Biomarkers: Time Trends

Models test whether Time linearly or quadratically predicts each biomarker. Quadratic was the better fit for most analytes.

## Model Comparison (AIC / BIC)

```{r model-comparison}
mlm_list <- list(
  CRP = mlm_crp, `IL-10` = mlm_il10, `IL-6` = mlm_il6, `TNF-alpha` = mlm_tnfalpha,
  DHEA = mlm_dhea, Testosterone = mlm_test, Cortisol = mlm_cort
)

model_comp_tbl <- imap_dfr(mlm_list, function(res, name) {
  tibble(
    Analyte = name,
    Model   = c("Linear", "Quadratic"),
    df      = res$aic$df,
    AIC     = round(res$aic$AIC, 1),
    BIC     = round(res$bic$BIC, 1)
  )
})

best_rows <- model_comp_tbl |>
  group_by(Analyte) |>
  mutate(is_best = AIC == min(AIC)) |>
  ungroup() |>
  pull(is_best)

model_comp_tbl |>
  kable() |>
  kable_styling(bootstrap_options = "hover", full_width = FALSE) |>
  row_spec(which(best_rows), bold = TRUE, color = "#2166ac")
```

## Fixed Effects: Quadratic Models {.tabset}

```{r quad-fixed-effects, results='asis'}
iwalk(mlm_list, function(res, name) {
  cat("\n\n### ", name, "\n\n")
  tbl <- broom.mixed::tidy(res$quadratic, effects = "fixed") |>
    mutate(across(where(is.numeric), \(x) round(x, 3)))
  tbl |>
    mutate(p.value = cell_spec(p.value, bold = p.value < 0.05,
      color = if_else(p.value < 0.05, "#2166ac", "black"))) |>
    kable(escape = FALSE) |>
    kable_styling(bootstrap_options = "hover", full_width = FALSE) |>
    print()
})
```

## Quadratic Curves

```{r quad-plots, fig.width=10, fig.height=4}
cytokines_long <- sos_cytokines_panel |>
  pivot_longer(c(IL10_log, IL6_log, TNFalpha_log), names_to = "biomarker", values_to = "log_conc") |>
  mutate(biomarker = case_match(
    biomarker,
    "IL10_log"     ~ "IL-10",
    "IL6_log"      ~ "IL-6",
    "TNFalpha_log" ~ "TNF-\u03b1"
  ))

cytokines_quad_plot <- ggplot(cytokines_long, aes(Time, log_conc, color = biomarker)) +
  stat_smooth(method = "lm", formula = y ~ x + I(x^2), linewidth = 1,
              aes(fill = after_scale(colour)), alpha = 0.2) +
  scale_colour_viridis_d() +
  ylab("Concentration (log-transformed)") + xlab("Time point") + ggtitle("Cytokines") + theme_minimal()

hormones_long <- sos_hormones_panel |>
  pivot_longer(c(DHEA_log, Testosterone_log, Cortisol_log), names_to = "biomarker", values_to = "log_conc") |>
  mutate(biomarker = case_match(
    biomarker,
    "DHEA_log"         ~ "DHEA",
    "Testosterone_log" ~ "Testosterone",
    "Cortisol_log"     ~ "Cortisol"
  ))

hormones_quad_plot <- ggplot(hormones_long, aes(Time, log_conc, color = biomarker)) +
  stat_smooth(method = "lm", formula = y ~ x + I(x^2), linewidth = 1,
              aes(fill = after_scale(colour)), alpha = 0.2) +
  scale_colour_viridis_d() +
  ylab("Concentration (log-transformed)") + xlab("Time point") + ggtitle("Hormones") + theme_minimal()

ggarrange(cytokines_quad_plot, hormones_quad_plot, ncol = 2)
```

---

# Reactivity: Peak (T3/T4) vs. Baseline (T1)

Each person's peak is the higher of T3 or T4. Reactivity score = peak − T1 (positive = increase from baseline).

## Peak vs. T1 t-tests

```{r peak-ttests}
peak_ttest_tbl <- peak_tests_raw |>
  map(broom::tidy) |>
  list_rbind(names_to = "analyte") |>
  mutate(analyte = case_match(
    analyte,
    "crp" ~ "CRP", "il10" ~ "IL-10", "il6" ~ "IL-6", "tnfa" ~ "TNF-alpha",
    "dhea" ~ "DHEA", "test" ~ "Testosterone", "cort" ~ "Cortisol"
  )) |>
  select(analyte, estimate, statistic, parameter, p.value) |>
  mutate(across(where(is.numeric), \(x) round(x, 3)))

peak_ttest_tbl |>
  mutate(p.value = cell_spec(p.value, bold = p.value < 0.05,
    color = if_else(p.value < 0.05, "#2166ac", "black"))) |>
  kable(col.names = c("Analyte", "Mean diff (peak − T1)", "t", "df", "p (one-tailed)"), escape = FALSE) |>
  kable_styling(bootstrap_options = "hover", full_width = FALSE)
```

## FDR Correction

```{r fdr}
fdr_tbl <- tibble(
  group   = c(rep("Immune", 4), rep("Hormones", 3)),
  analyte = c("CRP", "IL-10", "IL-6", "TNF-alpha", "DHEA", "Testosterone", "Cortisol"),
  p_raw   = round(peak_praw, 6),
  p_fdr   = peak_pfdr
)

fdr_tbl |>
  mutate(p_fdr = cell_spec(round(p_fdr, 4), bold = p_fdr < 0.05,
    color = if_else(p_fdr < 0.05, "#2166ac", "black"))) |>
  kable(col.names = c("Group", "Analyte", "p (raw)", "p (FDR)"), escape = FALSE) |>
  kable_styling(bootstrap_options = "hover", full_width = FALSE)
```

## Immune Markers {.tabset}

### All (grid)

```{r immune-grid, fig.width=10, fig.height=9}
crp_plot  <- biomarker_plot(sos_crp_long,  "crp_log_conc",  "CRP (log-transformed)",        peak_pfdr["crp"])
il10_plot <- biomarker_plot(sos_il10_long, "il10_log_conc", "IL-10 (log-transformed)",       peak_pfdr["il10"])
il6_plot  <- biomarker_plot(sos_il6_long,  "il6_log_conc",  "IL-6 (log-transformed)",        peak_pfdr["il6"])
tnfa_plot <- biomarker_plot(sos_tnfa_long, "tnfa_log_conc", "TNF-\u03b1 (log-transformed)",  peak_pfdr["tnfa"])

ggarrange(crp_plot, il10_plot, il6_plot, tnfa_plot,
  labels = c("CRP", "IL-10", "IL-6", "TNF-alpha"), ncol = 2, nrow = 2)
```

### CRP

```{r crp-plot, fig.width=5, fig.height=5}
crp_plot
```

### IL-10

```{r il10-plot, fig.width=5, fig.height=5}
il10_plot
```

### IL-6

```{r il6-plot, fig.width=5, fig.height=5}
il6_plot
```

### TNF-alpha

```{r tnfa-plot, fig.width=5, fig.height=5}
tnfa_plot
```

## Hormones {.tabset}

### All (grid)

```{r hormone-grid, fig.width=12, fig.height=5}
dhea_plot <- biomarker_plot(sos_dhea_long, "dhea_log_conc", "DHEA (log-transformed)",          peak_pfdr["dhea"])
test_plot <- biomarker_plot(sos_test_long, "test_log_conc", "Testosterone (log-transformed)",   peak_pfdr["test"])
cort_plot <- biomarker_plot(sos_cort_long, "cort_log_conc", "Cortisol (log-transformed)",       peak_pfdr["cort"])

ggarrange(dhea_plot, test_plot, cort_plot,
  labels = c("DHEA", "Testosterone", "Cortisol"), ncol = 3, nrow = 1)
```

### DHEA

```{r dhea-plot, fig.width=5, fig.height=5}
dhea_plot
```

### Testosterone

```{r test-plot, fig.width=5, fig.height=5}
test_plot
```

### Cortisol

```{r cort-plot, fig.width=5, fig.height=5}
cort_plot
```

## Reactivity Descriptives

```{r reactivity-descriptives}
bind_rows(
  biomarkers_difference |> select(-SampleID) |> summarise(across(everything(), \(x) mean(x, na.rm = TRUE))) |> mutate(stat = "Mean"),
  biomarkers_difference |> select(-SampleID) |> summarise(across(everything(), \(x) sd(x,   na.rm = TRUE))) |> mutate(stat = "SD")
) |>
  select(stat, everything()) |>
  rename(
    ΔCRP           = crp_difference,
    `ΔIL-10`       = il10_difference,
    `ΔIL-6`        = il6_difference,
    `ΔTNF-α`       = tnfa_difference,
    ΔDHEA          = dhea_difference,
    ΔTestosterone  = test_difference,
    ΔCortisol      = cort_difference
  ) |>
  kable(digits = 3)
```

---

# Correlations: Reactivity Measures

HR and biomarker reactivity difference scores. *Subjective stress ratings will be added once data are available.*

```{r corr-matrix, fig.width=8, fig.height=7}
diff_matrix <- all_difference |> select(-SampleID) |> as.matrix()
colnames(diff_matrix) <- c(
  "ΔHR: Task vs. Rest", "ΔHR: SC vs. SD", "ΔHR: SC vs. Other", "ΔHR: SC vs. Rest",
  "ΔCRP", "ΔIL-10", "ΔIL-6", "ΔTNF-α", "ΔDHEA", "ΔTestosterone", "ΔCortisol"
)
difference_corrs <- cor(diff_matrix, use = "pairwise.complete.obs")

corrplot(
  difference_corrs,
  method = "number",
  tl.col = "black",
  col    = colorRampPalette(c("#FDE725FF", "#29AF7FFF", "#440154FF"))(100)
)
```

## Correlation p-values

```{r corr-pvalues}
difference_rcorrs <- rcorr(diff_matrix)

difference_rcorrs$P |>
  as.data.frame() |>
  mutate(across(everything(), \(x) round(x, 3))) |>
  mutate(across(everything(), \(x) {
    sig  <- !is.na(x) & x < 0.05
    xlab <- if_else(is.na(x), "—", as.character(x))
    cell_spec(xlab, bold = sig, color = if_else(sig, "#2166ac", "black"))
  })) |>
  kable(escape = FALSE) |>
  kable_styling(bootstrap_options = "hover")
```
