---
title: "cleaning TAG w2 immune data"
author: "cfm"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    df_print: paged
    highlight: tango
    theme: united
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
  pdf_document:
    toc: yes
---


### setting up markdown {.tabset}

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = TRUE,
	warning = TRUE
)

options(scipen=999)
```

### loading required packages

```{r load Required Packages, message=FALSE, warning=FALSE, include=FALSE}
library(pacman)
pacman::p_load(dplyr, lme4, nlme, readxl, ggplot2, tidyr, lubridate, psych, stringr, e1071, install = TRUE)
```

### setting root path

```{r root path}

cas_dir = "/Volumes/psych-cog/dsnlab/TAG/"
proj_root = "/Volumes/psych-cog/dsnlab/TAG/projects/immune/"
data_root = "/Volumes/psych-cog/dsnlab/TAG/data/Immune/"


```

# Creating intraassay CSV 
## manually copied over raw concentrations from each "Human CRP saliva plate #.csv" into one CSV

## This example is for multiplex assays of: salivary CRP(panel 1), IL-10, IL-6, TNFa (panel 2), and secretory IgA (panel 3) from Wave 3 of the University of Oregon TAG Study for a substudy K01MH111951 (michelle.byrne@gmail.com)

**FROM ORIGINAL SCRIPT**: Calculate intra-assay CVs (from MFI, not concentration) from the sample (duplicate) CVs. Note that the duplicate sample %CVs reported in the masterfile from the lab ARE calculated from the MFI. Also, the units are PERCENTAGES even though these are not indicated (so if the value is 5.39, that is 5.39%, or 0.0539, not 539%).

*NOTE*: Looking back at previous excel spreadsheets that calculated the intraassays, we do use the raw concentrations. 

```{r creating intraassay CSV}

cleaned_intra <- read.csv(paste0(data_root, "Wave2/W2_raw_immune_data_NOT_CLEANED.csv")) %>%
  mutate(tagid = na_if(tagid, "↓")) %>%
  fill(tagid) %>%
  mutate(across(c(CRP, IFNg, il10, il4, il6, tnf_alpha, IgA),
                ~ ifelse(grepl("^[<>]", .), NA, as.numeric(.)),
                .names = "num_{.col}")) %>%
  group_by(tagid) %>%
  summarise(
    across(starts_with("num_"),
           list(mean = ~mean(., na.rm = TRUE),
                sd = ~sd(., na.rm = TRUE)),
           .names = "{.col}_{.fn}"),
    plate = first(plate)
  ) %>%
  mutate(
    across(ends_with("_sd"), 
           ~ . / get(sub("_sd$", "_mean", cur_column())) * 100,
           .names = "{sub('_sd$', '_cv', .col)}")
  ) %>%
  ungroup()

write.csv(cleaned_intra, file = paste0(data_root, "Wave2/script_output/TAG_W2_immune_intraassays_ids.csv"))


intra_plate_avgs <- cleaned_intra %>%
  group_by(plate) %>%
  summarise(cv_crp = mean(num_CRP_cv, na.rm = TRUE),
            cv_IFNg = mean(num_IFNg_cv, na.rm = TRUE),
            cv_il10 = mean(num_il10_cv, na.rm = TRUE),
            cv_il4 = mean(num_il4_cv, na.rm = TRUE),
            cv_il6 = mean(num_il6_cv, na.rm = TRUE),
            cv_tnfalpha = mean(num_tnf_alpha_cv, na.rm = TRUE),
            cv_IgA = mean(num_IgA_cv, na.rm = TRUE))

write.csv(intra_plate_avgs, file = paste0(data_root, "Wave2/script_output/TAG_W2_immune_intraassay_plate_CVs.csv"))


intra_cvs <- intra_plate_avgs %>%
  summarise(cv_crp = mean(as.numeric(cv_crp), na.rm = TRUE),
            cv_IFNg = mean(as.numeric(cv_IFNg), na.rm = TRUE),
            cv_il10 = mean(as.numeric(cv_il10), na.rm = TRUE),
            cv_il4 = mean(as.numeric(cv_il4), na.rm = TRUE),
            cv_il6 = mean(as.numeric(cv_il6), na.rm = TRUE),
            cv_tnfalpha = mean(as.numeric(cv_tnfalpha), na.rm = TRUE),
            cv_IgA = mean(as.numeric(cv_IgA), na.rm = TRUE))

write.csv(intra_cvs, file = paste0(data_root, "Wave2/script_output/TAG_W2_immune_intraassay_CVs_avg"))

print(intra_plate_avgs)
print(intra_cvs)

```

# Steps for cleaning, tidying, and reporting on inflammatory markers.

**FROM ORIGINAL SCRIPT**: Calculate inter-assay CVs (from MFI, not concentration) from the high and low controls. Note: there are 5 plates but controls were reused for plates 4 & 5 (so only 3 plates for inter-assays) in WAVE 1 *note need to check for w3, in w2 the plates do not use the same controls

*NOTE*: using result, not MFI per previous wave's calculations; ASK ABOUT INVALID CONTROLS.. in previous waves, seems like we still used the value

```{r inter_cvs}

controls <- read.csv(paste0(data_root, "Wave2/TAG_W2_immune_highlowcontrols.csv")) %>% 
  mutate(
    high_num = ifelse(grepl("^[<>]", high), NA, as.numeric(high)),
    low_num  = ifelse(grepl("^[<>]", low), NA, as.numeric(low))
  )

print(controls)

## means within plates within markers

inter_plate_avgs <- controls %>%
  group_by(marker, plate) %>%
  summarise(
    high_mean = mean(high_num, na.rm = TRUE),
    low_mean = mean(low_num, na.rm = TRUE),
    .groups = "drop"
  )
  
inter_cvs <- inter_plate_avgs %>%
  group_by(marker) %>%
  summarise(
    high_cv = (sd(high_mean, na.rm = TRUE) / mean(high_mean, na.rm = TRUE)) * 100,
    low_cv = (sd(low_mean, na.rm = TRUE) / mean(low_mean, na.rm = TRUE)) * 100,
    inter_cv = mean(c(high_cv, low_cv), na.rm = TRUE),
    .groups = "drop"
  )

print(inter_cvs)

write.csv(inter_cvs, file = paste0(data_root, "Wave2/script_output/TAG_W2_immune_intercvs.csv"))

```

Load data and deal with out of range values (too high or low for assay to detect). 
IMPORTANT: CRP and SIgA Concentrations are in ng/ml. ALL OTHERS are in pg/ml.
We change right censored to max of distribution plus increment (i.e., sensitivity units of the assay) and left censored to lower limit of detection of that assay minus increment:
- in pg/ml, IFN-ɣ: 2.9; IL-10: 1.1; IL-12p70: 0.6; IL-2: 1.0; IL-4: 4.5; IL-6: 0.9; TNF-ɑ: 0.7
- in ng/ml: CRP: 0.001; SIgA: 0.40

**FROM ORIGINAL W1 SCRIPT**: Completely missing samples (i.e., no sample was given or nothing could be assayed) will be imputed after this cleaning process.


```{r}

raw_immune <- read.csv(paste0(data_root, "Wave2/W2_raw_immune_data_NOT_CLEANED.csv")) %>%
  mutate(tagid = na_if(tagid, "↓")) %>%
  fill(tagid) %>%
  mutate(across(c(CRP, IFNg, il10, il4, il6, tnf_alpha, IgA),
                ~ case_when(
                    grepl("^<", .) ~ "OOR<",
                    grepl("^>", .) ~ "OOR>",
                    TRUE ~ .
                  ))) %>%
  mutate(across(c(CRP, IFNg, il10, il4, il6, tnf_alpha, IgA),
                ~ ifelse(grepl("^(OOR<|OOR>)", .), NA, as.numeric(.)),
                .names = "num_{.col}"))

oor_counts <- raw_immune %>%
  summarise(across(c(CRP, IFNg, il10, il4, il6, tnf_alpha, IgA),
                   list(
                     oor_left = ~sum(grepl("^OOR<", .), na.rm = TRUE),
                     oor_right = ~sum(grepl("^OOR>", .), na.rm = TRUE)
                   ),
                   .names = "{.col}_{.fn}"))

print(oor_counts)

marker_max <- raw_immune %>%
  summarise(across(starts_with("num_"),
                   ~ max(., na.rm = TRUE),
                   .names = "{.col}_max"))

default_increment <- 0.01
crp_increment <- 0.0001

min_crp = 0.001/2
min_il10 = 1.1/2
min_il6 = 0.9/2
min_tnfa = 0.7/2
min_iga = 0.40/2
min_ifng = 2.9/2
min_il4 = 4.5/2

raw_immune <- raw_immune %>%
  mutate(
    CRP = ifelse(CRP == "OOR>", marker_max$num_CRP_max + crp_increment, CRP),
    IFNg = ifelse(IFNg == "OOR>", marker_max$num_IFNg_max + default_increment, IFNg),
    il10 = ifelse(il10 == "OOR>", marker_max$num_il10_max + default_increment, il10),
    il4 = ifelse(il4 == "OOR>", marker_max$num_il4_max + default_increment, il4),
    il6 = ifelse(il6 == "OOR>", marker_max$num_il6_max + default_increment, il6),
    tnf_alpha = ifelse(tnf_alpha == "OOR>", marker_max$num_tnf_alpha_max + default_increment, tnf_alpha),
    IgA = ifelse(IgA == "OOR>", marker_max$num_IgA_max + default_increment, IgA)
  ) %>% 
  mutate(
    CRP = ifelse(CRP == "OOR<", min_crp - crp_increment, CRP),
    IFNg = ifelse(IFNg == "OOR<", min_ifng - default_increment, IFNg),
    il10 = ifelse(il10 == "OOR<", min_il10 - default_increment, il10),
    il4 = ifelse(il4 == "OOR<", min_il4 - default_increment, il4),
    il6 = ifelse(il6 == "OOR<", min_il6 - default_increment, il6),
    tnf_alpha = ifelse(tnf_alpha == "OOR<", min_tnfa - default_increment, tnf_alpha),
    IgA = ifelse(IgA == "OOR<", min_iga - default_increment, IgA)
  ) %>% 
  mutate(across(c(CRP, IFNg, il10, il4, il6, tnf_alpha, IgA), as.numeric)) %>%
  separate(tagid, into = c("tagid", "sample_date"), sep = " (?=\\d{1,2}-\\d{1,2}-\\d{2})") %>% 
  mutate(tagid = gsub("-", "_", tagid)) %>% 
  select(-starts_with("num_"), -sample_date) %>% 
  group_by(tagid) %>% 
    summarise(
    win_IgA = mean(IgA, na.rm = TRUE),
    win_CRP = mean(CRP, na.rm = TRUE),
    win_IFNg = mean(IFNg, na.rm = TRUE),
    win_il10 = mean(il10, na.rm = TRUE),
    win_il4 = mean(il4, na.rm = TRUE),
    win_il6 = mean(il6, na.rm = TRUE),
    win_tnf_alpha = mean(tnf_alpha, na.rm = TRUE),
    .groups = "drop"
  )
  
write.csv(raw_immune, file = paste0(data_root, "Wave2/script_output/TAG_W2_immune_winsorized.csv"))

```

Correct your data for flow rate
#### Flow rate = *(Volume of saliva)/(time sample took in sec)*
#### Corrected concentration = *(observed concentration) x (Flow rate)*
##### CRP is *not* flow rate dependent
##### SIgA *is* flow rate dependent
##### For all the other markers, there is no evidence (yet) that it is flow rate dependent but not really any work done to investigate this. Will only correct for SIgA for now.

```{r}

covariates <- read.csv(paste0(data_root, "Wave2/TAG_W2_immune_covariates.csv"))
vol_ml = 1.5

full <- left_join(raw_immune, covariates, by = "tagid") %>% 
    mutate(
    minutes = floor(immune_time_to_collect),
    seconds = (immune_time_to_collect - minutes) * 100,
    duration = (minutes * 60) + seconds,
    flowrate = 1.5 / duration,
    win_IgA_fr_corrected = win_IgA * flowrate
  ) %>% 
  select(-win_IgA)

```

### CFM - STOPPED HERE !! 
Check normality, outliers, and skew stats in the raw concentrations
```{r check_norm}

full_long <- full %>%
  pivot_longer(
    cols = starts_with("win_"),
    names_to = "marker",
    values_to = "value"
  )

ggplot(full_long, aes(x = value)) +
  geom_histogram(bins = 30, fill = "steelblue", color = "white") +
  facet_wrap(~ marker, scales = "free") +
  labs(title = "winsorized", x = "marker", y = "value") +
  theme_minimal()

ggplot(full_long, aes(x = marker, y = value)) +
  geom_boxplot(fill = "tomato", outlier.shape = 21) +
  labs(title = "winsorized", x = "marker", y = "value") +
  theme_minimal()

normality_stats <- full_long %>%
  group_by(marker) %>%
  summarise(
    skewness = skewness(value, na.rm = TRUE),
    kurtosis = kurtosis(value, na.rm = TRUE),
    requires_log = abs(skewness) > 2 | abs(kurtosis) > 2,
    .groups = "drop"
  )

full_long <- full_long %>%
  left_join(normality_stats %>% select(marker, requires_log), by = "marker") %>%
  mutate(
    value_ln = ifelse(requires_log, log(value), value)  
  )

log_normality_stats <- full_long %>%
  group_by(marker) %>%
  summarise(
    log_skewness = skewness(value_ln, na.rm = TRUE),
    log_kurtosis = kurtosis(value_ln, na.rm = TRUE),
    .groups = "drop"
  )

final_normality_stats <- normality_stats %>%
  left_join(log_normality_stats, by = "marker") %>%
  mutate(
    log_skew_flag = ifelse(abs(log_skewness) > 2, "Outside [-2, 2]", "Acceptable"),
    log_kurtosis_flag = ifelse(abs(log_kurtosis) > 2, "Outside [-2, 2]", "Acceptable")
  )

```

Winsorize any remaining outliers that are +/- 3 SD (if any)

```{r}
lncrp_up_limit <- (final_lncrp_summary$mean + 3*(final_lncrp_summary$sd))
lncrp_up_limit
lncrp_lo_limit <- (final_lncrp_summary$mean - 3*(final_lncrp_summary$sd))
lncrp_lo_limit

lnil10_up_limit <- (final_lnil10_summary$mean + 3*(final_lnil10_summary$sd))
lnil10_up_limit
lnil10_lo_limit <- (final_lnil10_summary$mean - 3*(final_lnil10_summary$sd))
lnil10_lo_limit

lnil6_up_limit <- (final_lnil6_summary$mean + 3*(final_lnil6_summary$sd))
lnil6_up_limit
lnil6_lo_limit <- (final_lnil6_summary$mean - 3*(final_lnil6_summary$sd))
lnil6_lo_limit

lntnfa_up_limit <- (final_lntnfa_summary$mean + 3*(final_lntnfa_summary$sd))
lntnfa_up_limit
lntnfa_lo_limit <- (final_lntnfa_summary$mean - 3*(final_lntnfa_summary$sd))
lntnfa_lo_limit

lnsiga_up_limit <- (final_lnsiga_summary$mean + 3*(final_lnsiga_summary$sd))
lnsiga_up_limit
lnsiga_lo_limit <- (final_lnsiga_summary$mean - 3*(final_lnsiga_summary$sd))
lnsiga_lo_limit


tucrp_up_limit <- (final_tucrp_summary$mean + 3*(final_tucrp_summary$sd))
tucrp_up_limit
tucrp_lo_limit <- (final_tucrp_summary$mean - 3*(final_tucrp_summary$sd))
tucrp_lo_limit

tuil10_up_limit <- (final_tuil10_summary$mean + 3*(final_tuil10_summary$sd))
tuil10_up_limit
tuil10_lo_limit <- (final_tuil10_summary$mean - 3*(final_tuil10_summary$sd))
tuil10_lo_limit

tuil6_up_limit <- (final_tuil6_summary$mean + 3*(final_tuil6_summary$sd))
tuil6_up_limit
tuil6_lo_limit <- (final_tuil6_summary$mean - 3*(final_tuil6_summary$sd))
tuil6_lo_limit

tutnfa_up_limit <- (final_tutnfa_summary$mean + 3*(final_tutnfa_summary$sd))
tutnfa_up_limit
tutnfa_lo_limit <- (final_tutnfa_summary$mean - 3*(final_tutnfa_summary$sd))
tutnfa_lo_limit

tusiga_up_limit <- (final_tusiga_summary$mean + 3*(final_tusiga_summary$sd))
tusiga_up_limit
tusiga_lo_limit <- (final_tusiga_summary$mean - 3*(final_tusiga_summary$sd))
tusiga_lo_limit

# First check if there are any outliers (if not, no need to winsorize). Also, report these numbers. 
lncrp_num_right_outliers <- length(na.omit(tag_w3_immune$ln_crp [tag_w3_immune$ln_crp > lncrp_up_limit])) 
lncrp_num_right_outliers
lncrp_num_left_outliers <- length(na.omit(tag_w3_immune$ln_crp [tag_w3_immune$ln_crp < lncrp_lo_limit]))
lncrp_num_left_outliers

lnil10_num_right_outliers <- length(na.omit(tag_w3_immune$ln_il10 [tag_w3_immune$ln_il10 > lnil10_up_limit])) 
lnil10_num_right_outliers
lnil10_num_left_outliers <- length(na.omit(tag_w3_immune$ln_il10 [tag_w3_immune$ln_il10 < lnil10_lo_limit])) 
lnil10_num_left_outliers

lnil6_num_right_outliers <- length(na.omit(tag_w3_immune$ln_il6 [tag_w3_immune$ln_il6 > lnil6_up_limit])) 
lnil6_num_right_outliers
lnil6_num_left_outliers <- length(na.omit(tag_w3_immune$ln_il6 [tag_w3_immune$ln_il6 < lnil6_lo_limit]))
lnil6_num_left_outliers

lntnfa_num_right_outliers <- length(na.omit(tag_w3_immune$ln_tnfa [tag_w3_immune$ln_tnfa > lntnfa_up_limit])) 
lntnfa_num_right_outliers
lntnfa_num_left_outliers <- length(na.omit(tag_w3_immune$ln_tnfa [tag_w3_immune$ln_tnfa < lntnfa_lo_limit])) 
lntnfa_num_left_outliers 

lnsiga_num_right_outliers <- length(na.omit(tag_w3_immune$ln_siga [tag_w3_immune$ln_siga > lnsiga_up_limit]))
lnsiga_num_right_outliers 
lnsiga_num_left_outliers <- length(na.omit(tag_w3_immune$ln_siga [tag_w3_immune$ln_siga < lnsiga_lo_limit]))
lnsiga_num_left_outliers

tucrp_num_right_outliers <- length(na.omit(tag_w3_immune$tu_crp [tag_w3_immune$tu_crp > tucrp_up_limit])) 
tucrp_num_right_outliers
tucrp_num_left_outliers <- length(na.omit(tag_w3_immune$tu_crp [tag_w3_immune$tu_crp < tucrp_lo_limit]))
tucrp_num_left_outliers

tuil10_num_right_outliers <- length(na.omit(tag_w3_immune$tu_il10 [tag_w3_immune$tu_il10 > tuil10_up_limit])) 
tuil10_num_right_outliers
tuil10_num_left_outliers <- length(na.omit(tag_w3_immune$tu_il10 [tag_w3_immune$tu_il10 < tuil10_lo_limit])) 
tuil10_num_left_outliers

tuil6_num_right_outliers <- length(na.omit(tag_w3_immune$tu_il6 [tag_w3_immune$tu_il6 > tuil6_up_limit])) 
tuil6_num_right_outliers
tuil6_num_left_outliers <- length(na.omit(tag_w3_immune$tu_il6 [tag_w3_immune$tu_il6 < tuil6_lo_limit]))
tuil6_num_left_outliers

tutnfa_num_right_outliers <- length(na.omit(tag_w3_immune$tu_tnfa [tag_w3_immune$tu_tnfa > tutnfa_up_limit])) 
tutnfa_num_right_outliers
tutnfa_num_left_outliers <- length(na.omit(tag_w3_immune$tu_tnfa [tag_w3_immune$tu_tnfa < tutnfa_lo_limit])) 
tutnfa_num_left_outliers 

tusiga_num_right_outliers <- length(na.omit(tag_w3_immune$tu_siga [tag_w3_immune$tu_siga > tusiga_up_limit]))
tusiga_num_right_outliers 
tusiga_num_left_outliers <- length(na.omit(tag_w3_immune$tu_siga [tag_w3_immune$tu_siga < tusiga_lo_limit]))
tusiga_num_left_outliers
```

* Note: Update for W3. (From W1: In this data, tukey has 1 left outlier for IFN-g and 1 left outlier for IL-10. For ln, 6 right IFN-g, 1 left IL-10, 3 right IL-12, 2 right IL-4, 1 right TNF-a. Winsorize those.)
```{r}
library(DescTools)
tag_w3_immune$ifng_tu_w <- Winsorize(tag_w3_immune$tu_ifng, minval = tuifng_lo_limit, maxval = tuifng_up_limit, na.rm = TRUE)
tag_w3_immune$il10_tu_w <- Winsorize(tag_w3_immune$tu_il10, minval = tuil10_lo_limit, maxval = tuil10_up_limit, na.rm = TRUE)

tag_w3_immune$ifng_ln_w <- Winsorize(tag_w3_immune$ln_ifng, minval = lnifng_lo_limit, maxval = lnifng_up_limit, na.rm = TRUE)
tag_w3_immune$il10_ln_w <- Winsorize(tag_w3_immune$ln_il10, minval = lnil10_lo_limit, maxval = lnil10_up_limit, na.rm = TRUE)
tag_w3_immune$il12_ln_w <- Winsorize(tag_w3_immune$ln_il12, minval = lnil12_lo_limit, maxval = lnil12_up_limit, na.rm = TRUE)
tag_w3_immune$il4_ln_w <- Winsorize(tag_w3_immune$ln_il4, minval = lnil4_lo_limit, maxval = lnil4_up_limit, na.rm = TRUE)
tag_w3_immune$tnfa_ln_w <- Winsorize(tag_w3_immune$ln_tnfa, minval = lntnfa_lo_limit, maxval = lntnfa_up_limit, na.rm = TRUE)
```


Output final cleaned data
```{r}
tag_w3_immune$crp_tu_cleaned <- tag_w3_immune$tu_crp
tag_w3_immune$il10_tu_cleaned <- tag_w3_immune$il10_tu_w
tag_w3_immune$il6_tu_cleaned <- tag_w3_immune$tu_il6
tag_w3_immune$tnfa_tu_cleaned <- tag_w3_immune$tu_tnfa
tag_w3_immune$siga_tu_cleaned <- tag_w3_immune$tu_siga

tag_w3_immune$crp_ln_cleaned <- tag_w3_immune$ln_crp
tag_w3_immune$il10_ln_cleaned <- tag_w3_immune$il10_ln_w
tag_w3_immune$il6_ln_cleaned <- tag_w3_immune$ln_il6
tag_w3_immune$tnfa_ln_cleaned <- tag_w3_immune$tnfa_ln_w
tag_w3_immune$siga_ln_cleaned <- tag_w3_immune$ln_siga

write.csv(tag_w3_immune, file="TAG_W3_immune_cleaned.csv")
```

